---
title: "fluorescence_data_analysis"
author: "Álvaro Gómez Pérez"
date: "2025-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Libraries
```{r}
library(dplyr)
library(data.table)
library(tidyr)
library(ggplot2)
```


Set up
```{r}
# Set up environment
working_from = "charite"

if (working_from == "home") {
  base_dir = "/home/alvaro/MyStuff/TSSc/"
} else
  if (working_from == "charite") {
    base_dir = "C:/MyStuff/TSSC/"
  }
```


Load data
```{r}
# Fluorescence data
raw_1 <- as.data.frame(fread(paste(base_dir, "Data/Boone_lab/Fluorescence/raw_1.csv", sep = "")))
raw_2 <- as.data.frame(fread(paste(base_dir, "Data/Boone_lab/Fluorescence/raw_2.csv", sep = "")))
raw_3 <- as.data.frame(fread(paste(base_dir, "Data/Boone_lab/Fluorescence/raw_3.csv", sep = "")))

## Put together replicates
full_fluorescence <- rbind(rbind(raw_1, raw_2), raw_3)

# tRNA synthetases data
synthetases <- as.data.frame(fread(paste(base_dir, "Data/synthetases/trna_synthetases.csv", sep="")))
```





# 1. Some initial plots
## 1.1. With the ratio as normalized by them, directly
Histograms
```{r}
# Rep 1, all timepoints - full range
ggplot(data = raw_1, aes(x = `het:hom_mNG:mScar_ratio`)) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  labs(title = "Histogram of final ratio - all timepoints",
       subtitle = "Replicate 1")

# Rep 1, all timepoints - 0 to 15
ggplot(data = raw_1, aes(x = `het:hom_mNG:mScar_ratio`)) +
  geom_histogram(binwidth = 0.1, col = "black", fill = "grey") +
  theme_light() +
  coord_cartesian(xlim = c(0, 15)) +
  labs(title = "Histogram of final ratio - all timepoints",
       subtitle = "Replicate 1")

# Rep 1, all timepoints, log2
ggplot(data = raw_1, aes(x = log2(`het:hom_mNG:mScar_ratio`))) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  coord_cartesian(xlim = c(-12, 12)) +
  labs(title = "Histogram of final ratio - all timepoints",
       subtitle = "Replicate 1")

# Rep 1, time = 0, log2
ggplot(data = subset(raw_1, TimeAt37 == 0), aes(x = log2(`het:hom_mNG:mScar_ratio`))) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  coord_cartesian(xlim = c(-12, 12),
                  ylim = c(0, 500)) +
  labs(title = "Histogram of final ratio - 0h",
       subtitle = "Replicate 1")

# Rep 1, time = 12, log2
ggplot(data = subset(raw_1, TimeAt37 == 12), aes(x = log2(`het:hom_mNG:mScar_ratio`))) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  coord_cartesian(xlim = c(-12, 12),
                  ylim = c(0, 500)) +
  labs(title = "Histogram of final ratio - 12h",
       subtitle = "Replicate 1")
```

Boxplots - replicate 1
```{r}
# Final ratio, all timepoints
ggplot(data = raw_1, aes(x = as.factor(TimeAt37), y = log2(`het:hom_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio at all timepoints",
       subtitle = "Replicate 1")

# Heterozygous ratio, all timepoints
ggplot(data = subset(raw_1, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(`average_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(average_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the non-normalized ratio at all timepoints",
       subtitle = "Replicate 1")

# Heterozygous mNG, all timepoints
ggplot(data = subset(raw_1, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mNG), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(mNG)") +
  labs(title = "Boxplots for the mNG intensity",
       subtitle = "Replicate 1")

# Heterozygous mScarlet, all timepoints
ggplot(data = subset(raw_1, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mScar), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(mScarlet)") +
  labs(title = "Boxplots for the mScarlet intensity",
       subtitle = "Replicate 1")
```

Boxplots - all replicates - heterozygous
```{r}
# Final ratio, all timepoints
ggplot(data = full_fluorescence, aes(x = as.factor(TimeAt37), y = log2(`het:hom_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio at all timepoints",
       subtitle = "All replicates")

# Heterozygous ratio, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(`average_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(average_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the non-normalized ratio at all timepoints",
       subtitle = "All replicates")

# Heterozygous mNG, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mNG), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(mNG) intensity") +
  labs(title = "Boxplots for homozygous mNG at all timepoints",
       subtitle = "All replicates")

# Heterozygous mScarlet, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mScar), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(mScarlet intensity)") +
  labs(title = "Boxplots for homozygous mScarlet at all timepoints",
       subtitle = "All replicates")
```

Boxplots - all replicates - homozygous
```{r}
# Homozygous ratio, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Homozygous"), aes(x = as.factor(TimeAt37), y = log2(`average_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(average_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the homozygous ratio at all timepoints",
       subtitle = "All replicates, homozygous WT strains only")

# Homozygous mNG, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Homozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mNG), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(mNG) intensity") +
  labs(title = "Boxplots for homozygous mNG at all timepoints",
       subtitle = "All replicates, homozygous WT strains only")

# Homozygous mScarlet, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Homozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mScar), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(mScarlet intensity)") +
  labs(title = "Boxplots for homozygous mScarlet at all timepoints",
       subtitle = "All replicates, homozygous WT strains only")
```

Let me do some statistical testing because these boxplots don't seem to change along time 
```{r}
# Final ratio, 0h vs. 12h across all replicates
x <- full_fluorescence$`het:hom_mNG:mScar_ratio`[full_fluorescence$TimeAt37 == 0 & is.finite(full_fluorescence$`het:hom_mNG:mScar_ratio`)]
y <- full_fluorescence$`het:hom_mNG:mScar_ratio`[full_fluorescence$TimeAt37 == 12 & is.finite(full_fluorescence$`het:hom_mNG:mScar_ratio`)]

test <- wilcox.test(x, y)
test
median(x, na.rm = T) - median(y, na.rm = T)



```
--> So, there is statistical significance for the 0h and 12h timepoints being different, even though the boxplots don't show it. I guess this is partly because we have so many data points? At least it should not be due to the super large values in at least one of the groups, since WMW is rank based, right? So that is something. Still, I don't know how to evaluate the effect size of the difference between the 2 groups, does the difference between medians make sense? Because the difference between means is so biased by these huge values at 0h


## 1.2. Repeat some of these plots using the ratio normalized to time 0
First of all look at the distribution of this normalized ratio and fix it if necessary - remove those with too high values for the ratio normalized to time 0, this should be below 1 in heterozygous strains, and around 1 in homozygous ones! Regardless of the temperature. 
```{r}
ggplot(data = full_fluorescence, aes(x = log2(ratio_norm_to_t0))) +
  geom_histogram() +
  theme_light()

# For now let's remove any with values over 3 - this is what Johannes talked about that most of these values are likely pure fantasy, from the machine making wrong measurements and then the ratios and corrections expanding on these errors
full_fluorescence <- full_fluorescence %>%
  filter(ratio_norm_to_t0 <= 3)
```

Boxplots - all replicates - heterozygous
```{r}
# Final ratio, all timepoints
ggplot(data = full_fluorescence, aes(x = as.factor(TimeAt37), y = log2(`het:hom_mNG:mScar_ratio`), fill = as.factor(`het:hom_mNG:mScar_ratio_status`))) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio at all timepoints",
       subtitle = "All replicates",
       fill = "Strain type")

# Final ratio noramlized to time 0, all timepoints
ggplot(data = full_fluorescence, aes(x = as.factor(TimeAt37), y = ratio_norm_to_t0, fill = as.factor(`het:hom_mNG:mScar_ratio_status`))) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio normalized to T0",
       subtitle = "All replicates, at all timepoints",
       fill = "Strain type") +
  coord_cartesian(ylim = c(-0.1, 3.1))

# Heterozygous ratio, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(`average_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(average_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the non-normalized ratio at all timepoints",
       subtitle = "All replicates")

# Heterozygous mNG, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mNG), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(mNG) intensity") +
  labs(title = "Boxplots for homozygous mNG at all timepoints",
       subtitle = "All replicates")

# Heterozygous mScarlet, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mScar), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(mScarlet intensity)") +
  labs(title = "Boxplots for homozygous mScarlet at all timepoints",
       subtitle = "All replicates")
```







# 2. Check if total_cell_count or undetectable_cells explain the ratio being constant
Boxplots
```{r}
# Total cell count, all timepoints
ggplot(data = full_fluorescence, aes(x = as.factor(TimeAt37), y = log2(total_cell_count), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(Total cell count)") +
  labs(title = "Boxplots for the total cell count at all timepoints",
       subtitle = "All replicates")

# Total cell count, all timepoints - group by homozygous or heterozygous
ggplot(data = full_fluorescence, aes(x = as.factor(TimeAt37), y = log2(total_cell_count), fill = as.factor(Status))) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(Total cell count)") +
  labs(title = "Boxplots for the total cell count at all timepoints",
       subtitle = "All replicates",
       fill = "Strain type")

# Undetectable cells mNG, all timepoints
ggplot(data = full_fluorescence, aes(x = as.factor(TimeAt37), y = log2(undetectable_cells_mNG), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(Undetectable cells mNG)") +
  labs(title = "Boxplots for undetectable cells mNG at all timepoints",
       subtitle = "All replicates")

# Undetectable cells mScar, all timepoints
ggplot(data = full_fluorescence, aes(x = as.factor(TimeAt37), y = log2(undetectable_cells_mScar), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(Undetectable cells mScar)") +
  labs(title = "Boxplots for undetectable cells mScar at all timepoints",
       subtitle = "All replicates")
```

Correlations
```{r}
# Final ratio vs. total cell count
ggplot(data = full_fluorescence, aes(x = log2(total_cell_count), y = log2(`het:hom_mNG:mScar_ratio`), col = TimeAt37)) +
  geom_point() +
  theme_light() +
  coord_cartesian(ylim = c(-25, 25))

# Final ratio vs. undetectable cells mNG
ggplot(data = full_fluorescence, aes(x = log2(undetectable_cells_mNG), y = log2(`het:hom_mNG:mScar_ratio`), col = TimeAt37)) +
  geom_point() +
  theme_light()

# Final ratio vs. undetectable cells mScar
ggplot(data = full_fluorescence, aes(x = log2(undetectable_cells_mScar), y = log2(`het:hom_mNG:mScar_ratio`), col = TimeAt37)) +
  geom_point() +
  theme_light()
```



# 3. Check tRNA synthetase ts-allele strains
```{r}
# Isolate fluorescence data for the synthetases and check for how many synthetases we have it
fluorescence_synthetases <- full_fluorescence %>%
  filter(Gene %in% synthetases$Gene.symbol)
length(unique(fluorescence_synthetases$Gene))


# Take the average across replicates
test <- fluorescence_synthetases %>%
  filter(`het:hom_mNG:mScar_ratio_status` == "mutant") %>%
  group_by(Gene, Status, TimeAt37) %>%
  summarise_at(vars(`het:hom_mNG:mScar_ratio`), list(average_ratio_across_replicates = mean))


# Get spaghetti plots
## Basic one
ggplot(data = test, aes(x = TimeAt37, y = average_ratio_across_replicates, col = Gene)) +
  geom_line(linewidth = 1) +
  theme_light() +
  labs(title = "Normalized mNG/mScarlet for TS tRNA synthetase strains") +
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)")
  

## Zoom in to see the ones at the bottom
ggplot(data = test, aes(x = TimeAt37, y = average_ratio_across_replicates, col = Gene)) +
  geom_line(linewidth = 1) +
  theme_light() +
  coord_cartesian(ylim = c(0, 1.65)) +
  labs(title = "Normalized mNG/mScarlet for TS tRNA synthetase strains",
       subtitle = "Zoomed in") +
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)")

# Get a subset of the synthetase dataset for these ones, for a slide
synths_temp <- synthetases %>%
  filter(Gene.symbol %in% full_fluorescence$Gene) %>%
  dplyr::select(Gene.name, Gene.symbol, Gene.secondaryIdentifier, mitochondrial, type, cofactor_or_enzyme)
```



# 4. Filter strains, keeping those which follow the pattern we want (decreasing ratio)
Filtering
```{r}
fluorescence_temp <- full_fluorescence %>%
  filter(Status == "Heterozygous")
strains <- unique(fluorescence_temp$Gene)

strains_with_right_pattern <- c()

for (gene in strains) {
  temp <- fluorescence_temp %>%
    filter(Gene == gene)
  
  if (length(unique(temp$TimeAt37)) == 7) {
    if ((mean(temp$`het:hom_mNG:mScar_ratio`[temp$TimeAt37 == 0], na.rm = T) >= mean(temp$`het:hom_mNG:mScar_ratio`[temp$TimeAt37 == 4], na.rm = T)) &
        (mean(temp$`het:hom_mNG:mScar_ratio`[temp$TimeAt37 == 4], na.rm = T) >= mean(temp$`het:hom_mNG:mScar_ratio`[temp$TimeAt37 == 8], na.rm = T)) &
        (mean(temp$`het:hom_mNG:mScar_ratio`[temp$TimeAt37 == 8], na.rm = T) >= mean(temp$`het:hom_mNG:mScar_ratio`[temp$TimeAt37 == 12], na.rm = T))) {
      strains_with_right_pattern <- c(strains_with_right_pattern, unique(temp$Gene))
    }
  }
}

fluorescence_filtered <- full_fluorescence %>%
  filter(Gene %in% strains_with_right_pattern)
```

Plots
```{r}
# Final ratio, all timepoints
ggplot(data = fluorescence_filtered, aes(x = as.factor(TimeAt37), y = log2(`het:hom_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio at all timepoints",
       subtitle = "All replicates")

# Spaghetti plot with these selected strains - need to average across replicates first - NEED TO FIX THIS
fluorescence_filtered <- fluorescence_filtered %>%
  group_by(Gene, Status, TimeAt37) %>%
  summarise_at(vars(`het:hom_mNG:mScar_ratio`), list(average_ratio_across_replicates = mean))

ggplot(data = fluorescence_filtered, aes(x = TimeAt37, y = average_ratio_across_replicates, col = Gene)) +
  geom_line(linewidth = 1) +
  theme_light() +
  labs(title = "Normalized mNG/mScarlet for selected strains") +
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)")
```



```{r}
check <- full_fluorescence %>%
  filter(TimeAt37 == 0)
```

















Quick check tRNA_KOs ODs - delete - or cut and paste to the appropriate file in that project
```{r}
ODs <- as.data.frame(fread("C:/MyStuff/tRNA_KOs/Data/OD_processed.csv"))

ggplot(data = ODs, aes(x = OD_processed)) +
  geom_histogram(binwidth = 0.05, col = "black", fill = "grey") +
  theme_light() +
  xlab("OD") +
  ylab("Count")
```


























