---
title: "fluorescence_data_analysis"
author: "Álvaro Gómez Pérez"
date: "2025-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Libraries
```{r}
library(dplyr)
library(data.table)
library(tidyr)
library(ggplot2)
```


Set up
```{r}
# Set up environment
working_from = "charite"

if (working_from == "home") {
  base_dir = "/home/alvaro/MyStuff/TSSc/"
} else
  if (working_from == "charite") {
    base_dir = "C:/MyStuff/TSSC/"
  }
```


Load data
```{r}
raw_1 <- as.data.frame(fread(paste(base_dir, "Data/Boone_lab/Fluorescence/raw_1.csv", sep = "")))
raw_2 <- as.data.frame(fread(paste(base_dir, "Data/Boone_lab/Fluorescence/raw_2.csv", sep = "")))
raw_3 <- as.data.frame(fread(paste(base_dir, "Data/Boone_lab/Fluorescence/raw_3.csv", sep = "")))

# Put together replicates
full_fluorescence <- rbind(rbind(raw_1, raw_2), raw_3)
```





Some initial naive plots, I'll organize this afterwards

Histograms
```{r}
# Rep 1, all timepoints - full range
ggplot(data = raw_1, aes(x = `het:hom_mNG:mScar_ratio`)) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  labs(title = "Histogram of final ratio - all timepoints",
       subtitle = "Replicate 1")

# Rep 1, all timepoints - 0 to 15
ggplot(data = raw_1, aes(x = `het:hom_mNG:mScar_ratio`)) +
  geom_histogram(binwidth = 0.1, col = "black", fill = "grey") +
  theme_light() +
  coord_cartesian(xlim = c(0, 15)) +
  labs(title = "Histogram of final ratio - all timepoints",
       subtitle = "Replicate 1")

# Rep 1, all timepoints, log2
ggplot(data = raw_1, aes(x = log2(`het:hom_mNG:mScar_ratio`))) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  coord_cartesian(xlim = c(-12, 12)) +
  labs(title = "Histogram of final ratio - all timepoints",
       subtitle = "Replicate 1")

# Rep 1, time = 0, log2
ggplot(data = subset(raw_1, TimeAt37 == 0), aes(x = log2(`het:hom_mNG:mScar_ratio`))) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  coord_cartesian(xlim = c(-12, 12),
                  ylim = c(0, 500)) +
  labs(title = "Histogram of final ratio - 0h",
       subtitle = "Replicate 1")

# Rep 1, time = 12, log2
ggplot(data = subset(raw_1, TimeAt37 == 12), aes(x = log2(`het:hom_mNG:mScar_ratio`))) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  coord_cartesian(xlim = c(-12, 12),
                  ylim = c(0, 500)) +
  labs(title = "Histogram of final ratio - 12h",
       subtitle = "Replicate 1")
```



Boxplots - replicate 1
```{r}
# Final ratio, all timepoints
ggplot(data = raw_1, aes(x = as.factor(TimeAt37), y = log2(`het:hom_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio at all timepoints",
       subtitle = "Replicate 1")

# Heterozygous ratio, all timepoints
ggplot(data = subset(raw_1, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(`average_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio at all timepoints",
       subtitle = "Replicate 1")

# Heterozygous mNG, all timepoints
ggplot(data = subset(raw_1, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mNG), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio at all timepoints",
       subtitle = "Replicate 1")
```

Boxplots - all replicates
```{r}
# Final ratio, all timepoints
ggplot(data = full_fluorescence, aes(x = as.factor(TimeAt37), y = log2(`het:hom_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("log2(het:hom_mNG:mScar_ratio)") +
  labs(title = "Boxplots for the final ratio at all timepoints",
       subtitle = "All replicates")

# Heterozygous ratio, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(`average_mNG:mScar_ratio`), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("average_mNG:mScar_ratio") +
  labs(title = "Boxplots for the heterozygous ratio at all timepoints",
       subtitle = "All replicates")

# Heterozygous mNG, all timepoints
ggplot(data = subset(full_fluorescence, Status == "Heterozygous"), aes(x = as.factor(TimeAt37), y = log2(average_intensity_mNG), group = TimeAt37)) +
  geom_boxplot() +
  theme_light() + 
  xlab("Time at 37ºC") +
  ylab("Average mNG intensity") +
  labs(title = "Boxplots for heterozygous mNG at all timepoints",
       subtitle = "All replicates")
```


Let me do some statistical testing because these boxplots don't seem to change along time 
```{r}
# Final ratio, 0h vs. 12h across all replicates
x <- full_fluorescence$`het:hom_mNG:mScar_ratio`[full_fluorescence$TimeAt37 == 0 & is.finite(full_fluorescence$`het:hom_mNG:mScar_ratio`)]
y <- full_fluorescence$`het:hom_mNG:mScar_ratio`[full_fluorescence$TimeAt37 == 12 & is.finite(full_fluorescence$`het:hom_mNG:mScar_ratio`)]

test <- wilcox.test(x, y)
test
median(x, na.rm = T) - median(y, na.rm = T)



```
--> So, there is statistical significance for the 0h and 12h timepoints being different, even though the boxplots don't show it. I guess this is partly because we have so many data points? At least it should not be due to the super large values in at least one of the groups, since WMW is rank based, right? So that is something. Still, I don't know how to evaluate the effect size of the difference between the 2 groups, does the difference between medians make sense? Because the difference between means is so biased by these huge values at 0h

























Quick check tRNA_KOs ODs - delete - or cut and paste to the appropriate file in that project
```{r}
ODs <- as.data.frame(fread("C:/MyStuff/tRNA_KOs/Data/OD_processed.csv"))

ggplot(data = ODs, aes(x = OD_processed)) +
  geom_histogram(binwidth = 0.05, col = "black", fill = "grey") +
  theme_light() +
  xlab("OD") +
  ylab("Count")
```


























