---
title: "Initial_growth_exploration"
output: html_document
date: "2025-01-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



We received some preliminary growth data for the strains in the library at 22ºC and 26ºC, both as AUC of growth curves and as raw data. 
Here I analyze both of them and compare the results. The goal of this was to figure out which temperatures would be optimal for the culture. 






-------------------------------------------------------- With their pre-processed data --------------------------------------------------------


# 0. Set up
## 0.1. Load libraries
```{r}
library(xlsx)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggrepel)
library(stringr)
```

## 0.2. Load data
```{r}
growth_agar <- as.data.frame(fread("C:/MyStuff/TSSC/Data/Growth/My versions/Processed_by_them/growth_agar.tsv"))
growth_liquid <- as.data.frame(fread("C:/MyStuff/TSSC/Data/Growth/My versions/Processed_by_them/growth_liquid.tsv"))
```



# 1. Start exploring
## 1.1. Some safety checks
```{r}
# Check rows annotated to be excluded in "SB.notes" column - in agar dataset
levels(as.factor(growth_agar$SB.notes))
bool_vect_exclude <- grepl("exclude", growth_agar$SB.notes, ignore.case = T)
bool_vect_exclude_empty <- grepl("empty", growth_agar$SB.notes)
sum(bool_vect_exclude)                                                  # 20 rows to be excluded
sum(bool_vect_exclude_empty)                                            # Out of them, 3 is because they are empty

# Check control strain - in agar dataset
bool_vect_control_strain <- grepl("control", growth_agar$SB.notes)
sum(bool_vect_control_strain)                                           # Only 1 row of control - makes sense since replicates are in columns

# Check other Notes column - in agar dataset
#table(as.factor(growth_agar$Notes))
      # All those that say "Repeat at ..."                  --> also "exclude" in the SB.notes
      # 1 "failed construction"                             --> also "exclude"
      # 3 "ignore colony - Li isn't sure what it is"        --> also "exclude"
      # 8 "Li's control BY4741 + p13623"                    --> also "exclude", not sure why but okay?
      # "P01-C12: apc11-22_Y11147" and "P01-C12: apc11-22_Y11147", 3rd try --> same IDs! Tag them manually as repeats of the same?
      # Many individual ones with structure: P01-D12: cdc36-16_Y08308, all other columns in order, not sure what this means
      # Wild type control, BY4741
      # Y1239 + p13623 control added by Fang                --> still not sure why this is the control strain but okay. Is it the strain they started from to add the different ts alleles?

# Remove unnecessary variables
rm(bool_vect_control_strain, bool_vect_exclude, bool_vect_exclude_empty)
```


## 1.2. Check if Notes are the same between agar and liquid datasets, identify problematic rows
```{r}
# Check that the levels of the Notes column are the same between datasets
length(levels(as.factor(growth_agar$Notes))) ==  length(levels(as.factor(growth_liquid$Notes)))                                     # Same number of levels
sum(levels(as.factor(growth_agar$Notes)) == levels(as.factor(growth_liquid$Notes)))/length(levels(as.factor(growth_agar$Notes)))    # Levels are exactly the same

# Check that the notes are exactly the same in the liquid and agar data for every row
liquid_temp <- growth_liquid %>%
  mutate(ID = paste(Sys.Name, Std.Name, TS.allele, Lab.ID, sep = "_")) %>%
  filter(Dilution == "1/10" & ID != "___") %>%
  select(Sys.Name, Std.Name, TS.allele, Lab.ID, Notes, SB.notes, ID) %>%
  rename(Notes_liquid = Notes,
         SB.notes_liquid = SB.notes)
agar_temp <- growth_agar %>%
  mutate(ID = paste(Sys.Name, Std.Name, TS.allele, Lab.ID, sep = "_")) %>%
  filter(ID != "___") %>%
  select(Sys.Name, Std.Name, TS.allele, Lab.ID, Notes, SB.notes, ID) %>%
  rename(Notes_agar = Notes,
         SB.notes_agar = SB.notes)
  
temp <- full_join(liquid_temp, agar_temp, by = c("ID"))                         # This is where the many-to-many relationship happens
sum(temp$Notes_liquid == temp$Notes_agar)
sum(temp$SB.notes_liquid == temp$SB.notes_agar)

## Pretty sure that notes are exactly the same between datasets - there are just some problematic rows because their IDs are repeated within each of the datasets
count_liquid_ID_reps <- c()
for (i in 1:nrow(liquid_temp)) {
  ID_liquid <- liquid_temp$ID[i]
  count_liquid_ID_reps <- c(count_liquid_ID_reps, sum(agar_temp$ID == ID_liquid))
}
problematic_rows_liquid <- liquid_temp[count_liquid_ID_reps != 1,]

count_agar_ID_reps <- c()
for (i in 1:nrow(agar_temp)) {
  ID_agar <- agar_temp$ID[i]
  count_agar_ID_reps <- c(count_agar_ID_reps, sum(liquid_temp$ID == ID_agar))
}
problematic_rows_agar <- agar_temp[count_agar_ID_reps != 1,]

## Check that these 2 dataframes are exactly equal to each other
ncol(problematic_rows_agar)*nrow(problematic_rows_agar) == (ncol(problematic_rows_liquid)*nrow(problematic_rows_liquid))    # Check same dimensions
sum(problematic_rows_agar == problematic_rows_liquid)/(ncol(problematic_rows_agar)*nrow(problematic_rows_agar))              # Check all position equal

## Sumarize all of this in a single dataframe and delete all the other stuff                                                    
problematic_rows <- problematic_rows_agar


# Remove unnecesary variables
rm(problematic_rows_agar, problematic_rows_liquid, agar_temp, liquid_temp, i, count_agar_ID_reps, count_liquid_ID_reps, ID_agar, ID_liquid, temp)
```


## 1.3. Get final version of the data to work with
```{r}
# Exclude rows to be excluded
growth_agar <- growth_agar[!(grepl("exclude", growth_agar$SB.notes, ignore.case = T)),]
growth_liquid <- growth_liquid[!(grepl("exclude", growth_liquid$SB.notes, ignore.case = T)),]

# Separate liquid into 2 datasets - for the 2 different dilutions
growth_liquid_1_10 <- growth_liquid %>% filter(Dilution == "1/10")
growth_liquid_1_100 <- growth_liquid %>% filter(Dilution == "1/100")
growth_liquid_full <- growth_liquid
rm(growth_liquid)
```


## 1.4. Initial plots
### For agar data
```{r}
# Overlayed histograms for the 2 temperatures
ggplot(data = growth_agar) +
  geom_histogram(aes(x = Avg_AUC_22, fill = "22ºC"), alpha = 0.4, position="identity") +
  geom_histogram(aes(x = Avg_AUC_26, fill = "26ºC"), alpha = 0.4, position="identity") +
  theme_light() +
  ylab("Count") +
  xlab("Average AUC across biological replicates") +
  scale_fill_manual(name = "Temperature",
                     breaks = c("22ºC", "26ºC"),
                     values = c("22ºC"="lightblue", "26ºC"="orange"))

# Get a histogram of the difference
ggplot(data = growth_agar, aes(x = Avg_AUC_26 - Avg_AUC_22)) +
  geom_histogram(col = "black", fill = "grey") +
  theme_light() +
  ylab("Count") +
  xlab("Diff. in average AUC (26ºC - 22ºC)")


# Get a scatter plot of Avg_AUC_22 vs. Avg_AUC_26 (same info as the histogram, but I think this one is more informative)
## Get the ratio between the 2 to color by it
growth_agar <- growth_agar %>%
  mutate(Ratio_26_to_22 = as.factor(case_when(Avg_AUC_26/Avg_AUC_22 < 0.8 ~ "Below_0.8",
                                    TRUE ~ "Above_0.8")))
## Get the plot
ggplot(data = growth_agar, aes(x = Avg_AUC_22, y = Avg_AUC_26)) +
  geom_point() +
  geom_abline() +
  theme_light() +
  labs(col = "Ratio 26ºC/22ºC") +
  labs(title = "With average AUCs across replicates") +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")


# Get the same scatter plot but using the AUCs normalized based on the plate average - I assume this kind of corrects for batch effects? Or that is their idea?
## Get the ratio
growth_agar <- growth_agar %>%
  mutate(Ratio_26_to_22_norm_by_plate = as.factor(case_when(Relative_AUC_to_plate_mean_26/Relative_AUC_to_plate_mean_22 < 0.8 ~ "Below_0.8",
                                                  TRUE ~ "Above_0.8")))
## Get the plot
ggplot(data = growth_agar, aes(x = Relative_AUC_to_plate_mean_22, y = Relative_AUC_to_plate_mean_26)) +
  geom_point() +
  geom_abline() +
  labs(col = "Ratio 26ºC/22ºC") +
  theme_light() +
  labs(title = "With (average) AUCs normalized to plate mean") +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")


# Get the same scatter plot but using the AUCs normalized based on the WTs in each plate
## Get the ratio
growth_agar <- growth_agar %>%
  mutate(Ratio_26_to_22_norm_by_control = as.factor(case_when(Relative_AUC_to_control_mean_26/Relative_AUC_to_control_mean_22 < 0.8 ~ "Below_0.8",
                                                  TRUE ~ "Above_0.8")))
## Get the plot
ggplot(data = growth_agar, aes(x = Relative_AUC_to_control_mean_22, y = Relative_AUC_to_control_mean_26)) +
  geom_point() +
  geom_abline() +
  labs(col = "Ratio 26ºC/22ºC") +
  theme_light() +
  labs(title = "With (average) AUCs normalized to plate mean") +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")


# Same plot but color by plate just to be sure there isn't a huge effect of that - there is :(
ggplot(data = growth_agar, aes(x = Relative_AUC_to_control_mean_22, y = Relative_AUC_to_control_mean_26, col = as.factor(Plate384))) +
  geom_point() +
  geom_abline() +
  labs(col = "Plate") +
  labs(title = "With (average) AUCs normalized to control mean") +
  theme_light() +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")



# Oh no, all the ones that seem to grow better between 22ºC are from a certain 384-well plate - come up with plate_column and plate_row variables that I can plot based on
growth_agar <- growth_agar %>%
  mutate(Plate_Column = paste(Plate384, Col384, sep="_"),
         Plate_Row = paste(Plate384, Row384, sep="_"))

# Plot based on Plate_Column
ggplot(data = growth_agar, aes(x = Avg_AUC_22, y = Avg_AUC_26, col = as.factor(Plate_Column))) +
  geom_point() +
  geom_abline() +
  labs(col = "Ratio 26ºC/22ºC") +
  theme_light() +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC") +
  geom_text_repel(data = subset(growth_agar, Ratio_26_to_22 == "Below_0.8"), aes(x = Avg_AUC_22, y = Avg_AUC_26, col = as.factor(Plate_Column), label = Plate_Column)) +
  theme(legend.position = "none")

# Plot based on Plate_Row
ggplot(data = growth_agar, aes(x = Avg_AUC_22, y = Avg_AUC_26, col = as.factor(Plate_Row))) +
  geom_point() +
  geom_abline() +
  labs(col = "Ratio 26ºC/22ºC") +
  theme_light() +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC") +
  geom_text_repel(data = subset(growth_agar, Ratio_26_to_22 == "Below_0.8"), aes(x = Avg_AUC_22, y = Avg_AUC_26, col = as.factor(Plate_Row), label = Plate_Column)) +
  theme(legend.position = "none")
```

### For liquid data 1/10
```{r}
# Overlayed histograms for the 2 temperatures
ggplot(data = growth_liquid_1_10) +
  geom_histogram(aes(x = Avg_AUC_22, fill = "22ºC"), alpha = 0.4, position="identity") +
  geom_histogram(aes(x = Avg_AUC_26, fill = "26ºC"), alpha = 0.4, position="identity") +
  theme_light() +
  ylab("Count") +
  xlab("Average AUC across biological replicates") +
  scale_fill_manual(name = "Temperature",
                     breaks = c("22ºC", "26ºC"),
                     values = c("22ºC"="lightblue", "26ºC"="orange"))

# Overlayed histogram colored by Platenum
ggplot(data = growth_liquid_1_10) +
  geom_histogram(aes(x = Avg_AUC_22, fill = Platenum, group = Platenum), alpha = 0.4, position="identity") +
  geom_histogram(aes(x = Avg_AUC_26, fill = Platenum, group = Platenum), alpha = 0.4, position="identity") +
  theme_light() +
  ylab("Count") +
  xlab("Average AUC across biological replicates")

# Get a histogram of the difference
ggplot(data = growth_liquid_1_10, aes(x = Avg_AUC_26 - Avg_AUC_22)) +
  geom_histogram(col = "black", fill = "grey") +
  theme_light() +
  ylab("Count") +
  xlab("Diff. in average AUC (26ºC - 22ºC)")

# Get a scatter plot of Avg_AUC_22 vs. Avg_AUC_26 (same info as the histogram, but I think this one is more informative)
## Get the ratio between the 2 to color by it
growth_liquid_1_10 <- growth_liquid_1_10 %>%
  mutate(Ratio_26_to_22 = as.factor(case_when(Avg_AUC_26/Avg_AUC_22 < 0.8 ~ "Below_0.8",
                                    TRUE ~ "Above_0.8")))

## Get the plot
ggplot(data = growth_liquid_1_10, aes(x = Avg_AUC_22, y = Avg_AUC_26)) +
  geom_point() +
  geom_abline() +
  theme_light() +
  labs(col = "Ratio 26ºC/22ºC") +
  labs(title = "With average AUCs across replicates") +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")

# Get the same scatter plot but using the AUCs normalized based on the plate average - I assume this kind of corrects for batch effects? Or that is their idea?
## Get the ratio
growth_liquid_1_10 <- growth_liquid_1_10 %>%
  mutate(Ratio_26_to_22_norm_by_plate = as.factor(case_when(Relative_AUC_to_plate_mean_26/Relative_AUC_to_plate_mean_22 < 0.8 ~ "Below_0.8",
                                                  TRUE ~ "Above_0.8")))

## Get the plot
ggplot(data = growth_liquid_1_10, aes(x = Relative_AUC_to_plate_mean_22, y = Relative_AUC_to_plate_mean_26)) +
  geom_point() +
  geom_abline() +
  labs(col = "Ratio 26ºC/22ºC") +
  theme_light() +
  labs(title = "With (average) AUCs normalized to plate mean") +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")


# Color by Platenum (whatever that is) just to be sure there isn't a huge effect of that - there is :(
ggplot(data = growth_liquid_1_10, aes(x = Avg_AUC_22, y = Avg_AUC_26, col = as.factor(Platenum))) +
  geom_point() +
  geom_abline() +
  labs(col = "Plate") +
  labs(title = "With average AUCs across replicates") +
  theme_light() +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")

## Get the plot
ggplot(data = growth_liquid_1_10, aes(x = Relative_AUC_to_plate_mean_22, y = Relative_AUC_to_plate_mean_26, col = as.factor(Platenum))) +
  geom_point() +
  geom_abline() +
  labs(col = "Ratio 26ºC/22ºC") +
  theme_light() +
  labs(title = "With (average) AUCs normalized to plate mean") +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")
```

### For liquid data 1/100
```{r}
# Overlayed histograms for the 2 temperatures
ggplot(data = growth_liquid_1_100) +
  geom_histogram(aes(x = Avg_AUC_22, fill = "22ºC"), alpha = 0.4, position="identity") +
  geom_histogram(aes(x = Avg_AUC_26, fill = "26ºC"), alpha = 0.4, position="identity") +
  theme_light() +
  ylab("Count") +
  xlab("Average AUC across biological replicates") +
  scale_fill_manual(name = "Temperature",
                     breaks = c("22ºC", "26ºC"),
                     values = c("22ºC"="lightblue", "26ºC"="orange"))

# Get a histogram of the difference
ggplot(data = growth_liquid_1_100, aes(x = Avg_AUC_26 - Avg_AUC_22)) +
  geom_histogram(col = "black", fill = "grey") +
  theme_light() +
  ylab("Count") +
  xlab("Diff. in average AUC (26ºC - 22ºC)")

# Get a scatter plot of Avg_AUC_22 vs. Avg_AUC_26 (same info as the histogram, but I think this one is more informative)
## Get the ratio between the 2 to color by it
growth_liquid_1_100 <- growth_liquid_1_100 %>%
  mutate(Ratio_26_to_22 = as.factor(case_when(Avg_AUC_26/Avg_AUC_22 < 0.8 ~ "Below_0.8",
                                    TRUE ~ "Above_0.8")))

## Get the plot
ggplot(data = growth_liquid_1_100, aes(x = Avg_AUC_22, y = Avg_AUC_26)) +
  geom_point() +
  geom_abline() +
  theme_light() +
  labs(col = "Ratio 26ºC/22ºC") +
  labs(title = "With average AUCs across replicates") +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")

# Get the same scatter plot but using the AUCs normalized based on the plate average - I assume this kind of corrects for batch effects? Or that is their idea?
## Get the ratio
growth_liquid_1_100 <- growth_liquid_1_100 %>%
  mutate(Ratio_26_to_22_norm_by_plate = as.factor(case_when(Relative_AUC_to_plate_mean_26/Relative_AUC_to_plate_mean_22 < 0.8 ~ "Below_0.8",
                                                  TRUE ~ "Above_0.8")))

## Get the plot
## Get the plot
ggplot(data = growth_liquid_1_100, aes(x = Relative_AUC_to_plate_mean_22, y = Relative_AUC_to_plate_mean_26)) +
  geom_point() +
  geom_abline() +
  labs(col = "Ratio 26ºC/22ºC") +
  theme_light() +
  labs(title = "With (average) AUCs normalized to plate mean") +
  xlab("AUC at 22ºC") +
  ylab("AUC at 26ºC")
```


## 1.5. Identify strains growing less in 22ºC
```{r}
# Based on their growth curves
weird_strains <- growth_agar %>% filter(Ratio_26_to_22 == "Below_0.8")
sort(unique(weird_strains$Row384))
sort(unique(weird_strains$Col384))

# Based on the article from Li et al., 2011
li_1 <- fread("C:/MyStuff/TSSC/Data/From articles/Li_et_al_2011/Li_et_al_S1.csv")
li_1 <- li_1 %>%
  rename(TS.allele = ts.allele) %>%
  filter(sick_temp_lower == 26 | sick_temp_upper == 26 | lethal_temp_lower == 26 | lethal_temp_upper == 26)

# Check how many of the weird ones based on growth data are among those sick at 22ºC in the article - well fuck me
sum(weird_strains$TS.allele %in% li_1$TS.allele)
```


























-------------------------------------------------------- With their raw growth data --------------------------------------------------------


# 0. Set up
## 0.1. Load libraries
```{r}
library(xlsx)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggrepel)
library(stringr)
library(DescTools)
```

## 0.2. Load data and preprocess
```{r}
raw_agar <- as.data.frame(fread("C:/MyStuff/TSSC/Data/Growth/Original data/Raw/agar_raw_data.tsv"))

# Remove rows that are tagged as "exclude"
raw_agar <- raw_agar[!(grepl("exclude", raw_agar$SB.notes, ignore.case = T)),]

# Remove rows that don't have neither a "Lab.ID" nor a "Notes" (the controls don't have the first but do have the second)
raw_agar <- raw_agar %>%
  filter(!(Lab.ID == "" & Notes == "")) %>%
  select(-`Sys Name`)

# Create a unique ID column and 2 other ID columns, each based on the positions in each of the plates (Plate_Row_Col structure)
raw_agar <- raw_agar %>%
  mutate(Unique.ID = case_when(Lab.ID != "" ~ paste(Lab.ID, "Positions", plate384, row384, col384, plate1536, row1536, col1536, sep = "_"),
                               Lab.ID == "" ~ paste(Notes, "Positions", plate384, row384, col384, plate1536, row1536, col1536, sep = "_")),
         Plate384.ID = paste(plate384, row384, col384, sep = "_"),
         Plate1536.ID = paste(plate1536, row1536, col1536, sep ="_"))
```



# 1. Come up with the AUC values
Doing this in a very naïve way: not removing outliers from the growth curve before getting the AUC or anything like that, maybe I should do that? I am also ignoring the circularity, although from what Johannes said that seems alright. 
```{r}
# Iterate over the unique wells and calculate the AUC for each of them
wells <- unique(raw_agar$Unique.ID)
samples <- c()
replicate <- c()
temperatures <- c()
AUCs <- c()


for (i in 1:length(wells)) {
  # Define which sample we are evaluating in this iteration
  well <- wells[i]
  
  # Divide the dataset for this sample into 4, based on temperature and replicate
  temp_22_rep_1 <- raw_agar %>% filter(Unique.ID == well & temperature == 22 & replicate == 1)
  temp_26_rep_1 <- raw_agar %>% filter(Unique.ID == well & temperature == 26 & replicate == 1)
  temp_22_rep_2 <- raw_agar %>% filter(Unique.ID == well & temperature == 22 & replicate == 2)
  temp_26_rep_2 <- raw_agar %>% filter(Unique.ID == well & temperature == 26 & replicate == 2)
  
  # Calculate the AUC for each sub-dataset and save them to a vector
  AUCs <- c(AUCs, AUC(x = as.numeric(as.POSIXct(temp_22_rep_1$ts)), y = temp_22_rep_1$size, na.rm = T), 
            AUC(x = as.numeric(as.POSIXct(temp_26_rep_1$ts)), y = temp_26_rep_1$size, na.rm = T),
            AUC(x = as.numeric(as.POSIXct(temp_22_rep_2$ts)), y = temp_22_rep_2$size, na.rm = T), 
            AUC(x = as.numeric(as.POSIXct(temp_26_rep_2$ts)), y = temp_26_rep_2$size, na.rm = T))
  
  # Save all other data (sample name, replicate, temperature) to other vectors so we can put them all together into a final dataset
  samples <- c(samples, well, well, well, well)
  temperatures <- c(temperatures, 22, 26, 22, 26)
  replicate <- c(replicate, 1, 1, 2, 2)
}


# Build a dataset from the obtained data, and make it look like theirs did - colnames I want to end up with:
# Plate384, Row384, Col384, Sys.Name, Std.Name, TS.allele, Lab.ID, Notes, SB.notes, Rep_1_AUC_22, Rep_2_AUC_22, Avg_AUC_22, Relative_AUC_to_control_mean_22, Relative_AUC_to_plate_mean_22, Rep_1_AUC_26, Rep_2_AUC_26, Avg_AUC_26, Relative_AUC_to_control_mean_26, Relative_AUC_to_plate_mean_26
AUCs_temp <- data.frame(Unique.ID = samples, temperature = temperatures, AUCs, replicate)
raw_agar_temp_1 <- raw_agar %>%
  select(-c(size, circularity, ts)) %>%
  distinct(Unique.ID, temperature, replicate, .keep_all = T)

raw_agar_temp_2 <- raw_agar %>%
  select(Unique.ID, temperature, replicate, plate384, row384, col384) %>%
  distinct(Unique.ID, temperature, replicate, .keep_all = T)


## Separate by replicates, rename columns
## In this process I create an ID for the 384-well plate so that I can collapse by it
AUCs_rep_1_22 <- AUCs_temp %>% 
  filter(replicate == 1 & temperature == 22) %>%
  left_join(raw_agar_temp_2, by = c("Unique.ID", "temperature", "replicate")) %>%
  mutate(Plate384.ID = paste(plate384, row384, col384, sep = "_")) %>%
  group_by(Plate384.ID) %>%
  mutate(Rep_1_AUC_22 = mean(AUCs)) %>%
  ungroup() %>%
  select(-c(AUCs, plate384, row384, col384)) %>%
  distinct(Plate384.ID, .keep_all = T) %>%
  left_join(raw_agar_temp, by = c("Unique.ID", "temperature", "replicate")) %>%
  relocate(Rep_1_AUC_22, .after = last_col())
  
AUCs_rep_2_22 <- AUCs_temp %>% 
  filter(replicate == 2 & temperature == 22) %>%
  left_join(raw_agar_temp_2, by = c("Unique.ID", "temperature", "replicate")) %>%
  mutate(Plate384.ID = paste(plate384, row384, col384, sep = "_")) %>%
  group_by(Plate384.ID) %>%
  mutate(Rep_2_AUC_22 = mean(AUCs)) %>%
  ungroup() %>%
  select(-c(AUCs, plate384, row384, col384)) %>%
  distinct(Plate384.ID, .keep_all = T) %>%
  select(Plate384.ID, Rep_2_AUC_22)

AUCs_rep_1_26 <- AUCs_temp %>% 
  filter(replicate == 1 & temperature == 26) %>%
  left_join(raw_agar_temp_2, by = c("Unique.ID", "temperature", "replicate")) %>%
  mutate(Plate384.ID = paste(plate384, row384, col384, sep = "_")) %>%
  group_by(Plate384.ID) %>%
  mutate(Rep_1_AUC_26 = mean(AUCs)) %>%
  ungroup() %>%
  select(-c(AUCs, plate384, row384, col384)) %>%
  distinct(Plate384.ID, .keep_all = T) %>%
  select(Plate384.ID, Rep_1_AUC_26)
  
AUCs_rep_2_26 <- AUCs_temp %>% 
  filter(replicate == 2 & temperature == 26) %>%
  left_join(raw_agar_temp_2, by = c("Unique.ID", "temperature", "replicate")) %>%
  mutate(Plate384.ID = paste(plate384, row384, col384, sep = "_")) %>%
  group_by(Plate384.ID) %>%
  mutate(Rep_2_AUC_26 = mean(AUCs)) %>%
  ungroup() %>%
  select(-c(AUCs, plate384, row384, col384)) %>%
  distinct(Plate384.ID, .keep_all = T) %>%
  select(Plate384.ID, Rep_2_AUC_26)


## Join these separate datasets 
final_AUCs <- left_join(AUCs_rep_1_26, AUCs_rep_2_22, by = c("Plate384.ID"))
final_AUCs <- left_join(final_AUCs, AUCs_rep_2_26, by = c("Plate384.ID"))
final_AUCs <- left_join(AUCs_rep_1_22, final_AUCs, by = c("Plate384.ID"))

## Get some final columns to have the same information as in the original one
final_AUCs <- final_AUCs %>%
  select(-replicate) %>%
  mutate(Avg_AUC_22 = mean(Rep_1_AUC_22, Rep_2_AUC_22),
         Avg_AUC_26 = mean(Rep_1_AUC_26, Rep_2_AUC_26))
```


# 2. Plot and see if it looks anything like the previous one
Okay my units are pretty different from theirs, so I guess we did this in different ways, but still the plots look pretty similar to each other. 
```{r}
ggplot(data = final_AUCs, aes(x = Avg_AUC_22, y = Avg_AUC_26, col = as.factor(plate384))) +
  geom_point() +
  geom_abline() +
  theme_light() +
  labs(col = "384-well plate")

ggplot(data = final_AUCs, aes(x = Avg_AUC_22, y = Avg_AUC_26, col = as.factor(plate1536))) +
  geom_point() +
  geom_abline() +
  theme_light() +
  labs(col = "1536-well plate")
```























