---
title: "initial_growth_curves_tRNA_synthetases_Wenxi"
output: html_document
date: "2025-03-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Libraries
```{r}
library(xlsx)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggrepel)
library(stringr)
library(glue)
library(viridis)
library(platetools)
library(gridExtra)
library(roxygen2)
set.seed(2025)
```




# 0. Set up
## 0.1. Set up directory
```{r}
working_from = "charite"

if (working_from == "home") {
  base_dir = "/home/alvaro/MyStuff/"
} else
  if (working_from == "charite") {
    base_dir = "C:/MyStuff/"
  }


which_try = "second_try"                 # This indicates which of the 2 sets of OD measurements Wenxi did we want to look at 
if (which_try == "first_try") {
  try_for_plots = "First round"
} else
  if (which_try == "second_try") {
    try_for_plots = "Second round"
  }

```

## 0.2. Load data - NEED TO UPDATE THIS WHEN I GET THE REMAINING DATASETS
```{r}
if (which_try == "first_try") {
  # First try - data from the first time Wenxi did the growth curves - only 24h
  growth_26 <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/first_try/20250530-26C.xlsx", sep=""), sheetIndex = 1)
  growth_30 <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/first_try/20250529-30C.xlsx", sep=""), sheetIndex = 1)
  growth_34 <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/first_try/20250602-34C.xlsx", sep=""), sheetIndex = 1)
  growth_37 <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/first_try/20250602-37C.xlsx", sep=""), sheetIndex = 1)
} else 
  if (which_try == "second_try") {
    # Second try - data from the second time Wenxi did the growth curves - 48h
    growth_22 <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/second_try/20250623 TSGC-22-48H.xlsx", sep=""), sheetIndex = 1)
    growth_26 <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/second_try/20250606-TSGC-26C-48.xlsx", sep=""), sheetIndex = 1)
    growth_30 <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/second_try/20250623-TSGC-30C-48.xlsx", sep=""), sheetIndex = 1)
    growth_34 <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/second_try/20250606-TSGC-34C-48.xlsx", sep=""), sheetIndex = 1)
}


# Sample IDs (metadata to match sample names)
sample_ids <- read.xlsx(file = paste(base_dir, "TSSC/Data/pilot/pre_test_growth_curves/strains_for_pilot.xlsx", sep=""), sheetIndex = 1) %>%
  dplyr::select(GC96, Sys.Name, Std.Name, TS.allele, Lab.ID, Synthetase, Gene.name, Status, Lethal_temp) %>%
  dplyr::mutate(Sample.ID = paste(Sys.Name, "_1", sep=""))
```




# 1. Process growth data from PlateReader or whatever
## 1.1. Create functions to pre-process plate reader growth curves - TODO: NEED TO DOCUMENT THESE FUNCTIONS
```{r}
#' This function changes the sample names from the well ID to sample ID
#' 
function_to_set_new_colnames <- function(df, sample_ids) {
  old_colnames <- data.frame(colnames(df)[3:ncol(df)])
  colnames(old_colnames) <- c("GC96")
  temp_ids <- sample_ids %>%
    dplyr::select(GC96, Sample.ID)
  old_colnames <- left_join(old_colnames, temp_ids, by = "GC96")
  new_colnames <- c("Time", "Temperature", as.character(old_colnames$Sample.ID))
  
  colnames(df) <- new_colnames
  return(df)
}

#' Normalize the ODs for each sample to its OD at T = 0
#'
normalize_to_time_0 <- function(df) {
  keep_first_columns <- df %>%
    dplyr::select(Time, Temperature)
  only_data <- df %>%
    dplyr::select(!c(Time, Temperature))
  only_data <- as.data.frame(t(apply(as.matrix(only_data), 1, function(x) x/as.matrix(only_data)[1,])))
  
  out <- cbind(keep_first_columns, only_data)
  return(out)
}

#' Iterate over the strains, getting their averages and creating a new column for each of them
#' 
function_to_get_means_across_replicates <- function(df, strains) {
  # Come up with the average columns
  for (i in 1:length(strains)) {
    strain <- strains[i]
    df <- df %>%
      rowwise() %>%
      dplyr::mutate("{paste0(strain)}" := mean(c_across(contains(strain))), .after = Temperature)
  }
  
  # Remove all other columns
  df <- df %>%
    dplyr::select(Time, Temperature, any_of(strains))
  
  # Return the finished dataframe
  return(df)
}

#' Add a column with the temperature that they were supposed to have, then pivot them to long version
make_long_and_add_fixed_temp <- function(df, fixed_temp) {
  df <- df %>%
    dplyr::mutate(Temp_set = fixed_temp, .after = Temperature) %>%
    dplyr::rename(Temp_observed = Temperature) %>%
    pivot_longer(!c(Time, Temp_set, Temp_observed), names_to = "Strain", values_to = "OD_raw")
  return(df)
}




#' Wrapper function: takes the just imported data and returns the processed data, in either long or short format, normalized or not
#' 
#' @param df The data should be provided as a dataframe where the information of the machine and method used
#' are not included, only the OD measurements and the respective time and temperature
#' @param sample_ids Metadata about the samples, containing ...
#' @param replicates
#' @param empty_rows
#' @param fixed_temp
#'
process_plate_reader_growth_curves <- function(df, 
                                               sample_ids,
                                               replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                               empty_rows = c("D", "H"),
                                               fixed_temp, 
                                               normalized_to_t0, 
                                               long_version,
                                               wrong_plate) {
  
  # First of all we make sure we have the timepoints on the rows and the wells on the columns,
  # since one of the machines produces a dataframe in the opposite orientation
  if (sum(grepl("Cycle.Nr.", colnames(df))) == 1) {
    new_colnames <- df$Cycle.Nr.
    df <- df %>%
      dplyr::select(-Cycle.Nr.)
    df <- as.data.frame(t(df)) 
    colnames(df) <- new_colnames
  }
  
  # Set the first two colnames to be "Time" and "Temperature", also get rid of rownames
  colnames(df)[1] <- "Time"
  colnames(df)[2] <- "Temperature"
  rownames(df) <- NULL
  
  # Remove NAs in case full NA rows were introduced by Excel
  df <- na.omit(df)
  
  # There was a plate in which D row was not left empty (30ºC, first try) - need to do this horrible thing to take care of that
  if (wrong_plate == T) {
    ##   - row D in the plate is now row E in my dataset
    ##   - row E in the plate is now row F in my dataset
    ##   - row F in the plate is now row G in my dataset
    ##   - row G in the plate (empty) is now row D in my dataset
    ## SO REMEMBER THAT LABELS IN THE ANALYSIS FOR ROWS FROM D ARE CHANGED HERE!
    new_colnames <- c("Time", "Temperature",
                      paste("A", seq(1, 12, 1), sep = ""),
                      paste("B", seq(1, 12, 1), sep = ""), 
                      paste("C", seq(1, 12, 1), sep = ""), 
                      paste("E", seq(1, 12, 1), sep = ""), 
                      paste("F", seq(1, 12, 1), sep = ""),
                      paste("G", seq(1, 12, 1), sep = ""),
                      paste("D", seq(1, 12, 1), sep = ""),
                      paste("H", seq(1, 12, 1), sep = ""))
    colnames(df) <- new_colnames
  }
  
  # Extract the names of the rows that contain the replicates
  r1 <- replicates[[1]][1]
  r2 <- replicates[[1]][2]
  r3 <- replicates[[1]][3]
  r4 <- replicates[[2]][1]
  r5 <- replicates[[2]][2]
  r6 <- replicates[[2]][3]
  empty_1 <- empty_rows[1]
  empty_2 <- empty_rows[2]
  
  # Sample IDs - replicate it a couple times changing IDs, since the samples in wells A, B and C are triplicates of the same sample, and the same for E, F, G
  sample_ids_2 <- sample_ids %>%
    dplyr::mutate(GC96 = case_when(grepl(r1, GC96) ~ str_replace(GC96, r1, r2),
                                   grepl(r4, GC96) ~ str_replace(GC96, r4, r5)),
                  Sample.ID = paste(Sys.Name, "_2", sep=""))
  sample_ids_3 <- sample_ids %>%
    dplyr::mutate(GC96 = case_when(grepl(r1, GC96) ~ str_replace(GC96, r1, r3),
                                   grepl(r4, GC96) ~ str_replace(GC96, r4, r6)),
                  Sample.ID = paste(Sys.Name, "_3", sep=""))
  
  sample_ids_new <- rbind(rbind(sample_ids, sample_ids_2), sample_ids_3)
  
  # Need to add rows for the empty wells, otherwise I get an error later...
  d_sample_ids <- data.frame(c(paste(empty_1, seq(1, 12, 1), sep="")),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             paste("Empty_", seq(1, 12, 1), sep=""))
  colnames(d_sample_ids) <- colnames(sample_ids_new)
  
  h_sample_ids <- data.frame(c(paste(empty_2, seq(1, 12, 1), sep="")),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             rep("Empty", 12),
                             paste("Empty_", seq(13, 24, 1), sep=""))
  colnames(h_sample_ids) <- colnames(sample_ids_new)
  
  sample_ids_new_2 <- rbind(rbind(sample_ids_new, d_sample_ids), h_sample_ids)
  
  
  # Set new column names, take averages of the 3 replicates, set to long format and put all temperatures together --> ready to plot!
  ## Set new colnames
  df <- function_to_set_new_colnames(df, sample_ids_new_2)
  
  ## Normalize to time 0
  df_norm <- normalize_to_time_0(df)
  
  ## Get mean across replicates
  strains <- unique(sample_ids_new_2$Sys.Name)
  df_norm_avg <- function_to_get_means_across_replicates(df_norm, strains)
  df_avg <- function_to_get_means_across_replicates(df, strains)
  
  ## Set to long format and add reference temperature column
  df_norm_avg_long <- make_long_and_add_fixed_temp(df_norm_avg, fixed_temp)
  df_avg_long <- make_long_and_add_fixed_temp(df_avg, fixed_temp)
  
  # Figure out the required type of output, and return it
  if (normalized_to_t0 == T & long_version == T) {
    df_norm_avg_long <- df_norm_avg_long %>%
      dplyr::rename(OD_norm = OD_raw)
    return(df_norm_avg_long)
  }
  else if (normalized_to_t0 == T & long_version == F) {
    df_norm_avg <- df_norm_avg %>%
      dplyr::rename(OD_norm = OD_raw)
    return(df_norm_avg)
  }
  else if (normalized_to_t0 == F & long_version == T) {
    return(df_avg_long)
  }
  else if (normalized_to_t0 == F & long_version == F) {
    return(df_avg)
  }
}
```

## 1.2. Run pre-processing - for the datasets from the first set of growth curves
Here I apply all the functions created above to the datasets from both the first and second set of growth curves done by Wenxi. I must have separate code for each of these 2 options since there are different temperatures in the 2 sets, and the 30ºC one in the first set has the wrong row empty. 
```{r}
# Those from the first set
if (which_try == "first_try") {
  # Here we get: 
  # - Long version
  # - Normalized to time 0
  # - Averaged across replicates
  growth_26_avg_norm_long <- process_plate_reader_growth_curves(df = growth_26,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 26,
                                                           normalized_to_t0 = T,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_30_avg_norm_long <- process_plate_reader_growth_curves(df = growth_30,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 30,
                                                           normalized_to_t0 = T,
                                                           long_version = T,
                                                           wrong_plate = T)
  
  growth_34_avg_norm_long <- process_plate_reader_growth_curves(df = growth_34,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 34,
                                                           normalized_to_t0 = T,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_37_avg_norm_long <- process_plate_reader_growth_curves(df = growth_37,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 37,
                                                           normalized_to_t0 = T,
                                                           long_version = T,
                                                           wrong_plate = F)
  # Join results from this into a single long dataframe with all temperatures, to plot from
  growth_long_full_norm <- rbind(rbind(rbind(growth_26_avg_norm_long, growth_30_avg_norm_long), growth_34_avg_norm_long), growth_37_avg_norm_long)
  
  
  
  # Here we get:
  # - Long version
  # - Not normalized
  # - Averaged across replicates
  growth_26_avg_long_raw <- process_plate_reader_growth_curves(df = growth_26,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 26,
                                                           normalized_to_t0 = F,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_30_avg_long_raw <- process_plate_reader_growth_curves(df = growth_30,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 30,
                                                           normalized_to_t0 = F,
                                                           long_version = T,
                                                           wrong_plate = T)
  
  growth_34_avg_long_raw <- process_plate_reader_growth_curves(df = growth_34,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 34,
                                                           normalized_to_t0 = F,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_37_avg_long_raw <- process_plate_reader_growth_curves(df = growth_37,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 37,
                                                           normalized_to_t0 = F,
                                                           long_version = T,
                                                           wrong_plate = F)
  # Join results from this into a single long dataframe with all temperatures, to plot from
  growth_long_full_raw <- rbind(rbind(rbind(growth_26_avg_long_raw, growth_30_avg_long_raw), growth_34_avg_long_raw), growth_37_avg_long_raw)
}




# Those from the second set
if (which_try == "second_try") {
  # Here we get: 
  # - Long version
  # - Normalized to time 0
  # - Averaged across replicates
  growth_22_avg_norm_long <- process_plate_reader_growth_curves(df = growth_22,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 22,
                                                           normalized_to_t0 = T,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_26_avg_norm_long <- process_plate_reader_growth_curves(df = growth_26,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 26,
                                                           normalized_to_t0 = T,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_30_avg_norm_long <- process_plate_reader_growth_curves(df = growth_30,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 30,
                                                           normalized_to_t0 = T,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_34_avg_norm_long <- process_plate_reader_growth_curves(df = growth_34,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 34,
                                                           normalized_to_t0 = T,
                                                           long_version = T,
                                                           wrong_plate = F)
  # Join results from this into a single long dataframe with all temperatures, to plot from
  growth_long_full_norm <- rbind(rbind(rbind(growth_22_avg_norm_long, growth_26_avg_norm_long), growth_30_avg_norm_long), growth_34_avg_norm_long)
  
  
  # Here we get: 
  # - Wide version
  # - No normalization
  # - Averaged across replicates
  # --> In order to check what was the issue at 30ºC!
  growth_22_avg <- process_plate_reader_growth_curves(df = growth_22,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 22,
                                                           normalized_to_t0 = F,
                                                           long_version = F,
                                                           wrong_plate = F)
  
  growth_26_avg <- process_plate_reader_growth_curves(df = growth_26,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 26,
                                                           normalized_to_t0 = F,
                                                           long_version = F,
                                                           wrong_plate = F)
  
  growth_30_avg <- process_plate_reader_growth_curves(df = growth_30,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 30,
                                                           normalized_to_t0 = F,
                                                           long_version = F,
                                                           wrong_plate = F)
  
  growth_34_avg <- process_plate_reader_growth_curves(df = growth_34,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 34,
                                                           normalized_to_t0 = F,
                                                           long_version = F,
                                                           wrong_plate = F)
  
  
  # Here we get: 
  # - Long version
  # - No normalization
  # - Averaged across replicates
  growth_22_avg_long_raw <- process_plate_reader_growth_curves(df = growth_22,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 22,
                                                           normalized_to_t0 = F,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_26_avg_long_raw <- process_plate_reader_growth_curves(df = growth_26,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 26,
                                                           normalized_to_t0 = F,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_30_avg_long_raw <- process_plate_reader_growth_curves(df = growth_30,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 30,
                                                           normalized_to_t0 = F,
                                                           long_version = T,
                                                           wrong_plate = F)
  
  growth_34_avg_long_raw <- process_plate_reader_growth_curves(df = growth_34,
                                                           sample_ids = sample_ids, 
                                                           replicates = list(c("A", "B", "C"), c("E", "F", "G")),
                                                           empty_rows = c("D", "H"),
                                                           fixed_temp = 34,
                                                           normalized_to_t0 = F,
                                                           long_version = T,
                                                           wrong_plate = F)
  # Join results from this into a single long dataframe with all temperatures, to plot from
  growth_long_full_raw <- rbind(rbind(rbind(growth_22_avg_long_raw, growth_26_avg_long_raw), growth_30_avg_long_raw), growth_34_avg_long_raw)
}
```



# 2. Initial checks
## 2.1. How stable was the temperature along each of the cultures?
```{r}
# In the first set of measurements, first two temperatures were in one incubator, next two in another one, should check the names and put them somewhere in the graph - similar for the second set, there is a mix of the 2 different incubators/PlateReaders/whatever they are
# They should be in the second sheet of each of these Excel files
# Be careful, I am not doing this from the original data but from the long version, so if I change anything it might get biased (more observations from one temperature than another or whatever)

if (which_try == "first_try") {
  ggplot() +
  geom_line(data = subset(growth_long_full_norm, Temp_set == 26), aes(x = Time, y = Temp_observed), col = "darkgreen", linewidth = 1) +
  geom_line(data = subset(growth_long_full_norm, Temp_set == 30), aes(x = Time, y = Temp_observed), col = "blue", linewidth = 1) +
  geom_line(data = subset(growth_long_full_norm, Temp_set == 34), aes(x = Time, y = Temp_observed), col = "purple", linewidth = 1) +
  geom_line(data = subset(growth_long_full_norm, Temp_set == 37), aes(x = Time, y = Temp_observed), col = "orange", linewidth = 1) +
  geom_hline(yintercept = 26, col = "black") +
  geom_hline(yintercept = 30, col = "black") +
  geom_hline(yintercept = 34, col = "black") +
  geom_hline(yintercept = 37, col = "black") +
  theme_light() +
  xlab("Time (s)") +
  ylab("Temperature") +
  labs(title = "How well did the incubators maintain their assigned temperature?",
       subtitle = try_for_plots)
} else 
  if (which_try == "second_try") {
  ggplot() +
  geom_line(data = subset(growth_long_full_norm, Temp_set == 22), aes(x = Time, y = Temp_observed), col = "orange", linewidth = 1) +
  geom_line(data = subset(growth_long_full_norm, Temp_set == 26), aes(x = Time, y = Temp_observed), col = "darkgreen", linewidth = 1) +
  geom_line(data = subset(growth_long_full_norm, Temp_set == 30), aes(x = Time, y = Temp_observed), col = "blue", linewidth = 1) +
  geom_line(data = subset(growth_long_full_norm, Temp_set == 34), aes(x = Time, y = Temp_observed), col = "purple", linewidth = 1) +
  geom_hline(yintercept = 22, col = "black") +
  geom_hline(yintercept = 26, col = "black") +
  geom_hline(yintercept = 30, col = "black") +
  geom_hline(yintercept = 34, col = "black") +
  theme_light() +
  xlab("Time (s)") +
  ylab("Temperature") +
  labs(title = "How well did the incubators maintain their assigned temperature?",
       subtitle = try_for_plots)
}

```

## 2.2. How consistent is the initial OD across all samples and temperatures - to see if normalizing to it is reasonable
I would say that at least for the second round it does look pretty good?? All initial ODs are within a range of 0.025 OD, seems reasonable to me. 34ºC does look generally a bit higher than the other temperatures. 
```{r}
ggplot(data = subset(growth_long_full_raw, Time == 0), 
       aes(x = Strain, y = OD_raw, col = as.factor(Temp_set))) +
  geom_point() +
  theme_light() +
  labs(title = "ODs at time = 0, for each strain at each temperature",
       subtitle = try_for_plots,
       col = "Temperature") +
  ylab("OD") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

## 2.3. Check what is wrong with T = 30ºC in the second try of the pilot - no idea, but it is wrong lol, Wenxi is going to repeat it, I think I can delete this
```{r}
#t_26_just_data <- growth_26_avg %>%
#  dplyr::select(-c(Time, Temperature))
#t_30_just_data <- growth_30_avg %>%
#  dplyr::select(-c(Time, Temperature))
#t_34_just_data <- growth_34_avg %>%
#  dplyr::select(-c(Time, Temperature))
#
#
#hist(as.matrix(t_26_just_data))
#hist(as.matrix(t_30_just_data))
#hist(as.matrix(t_34_just_data))
```



# 3. Main growth curves
## 3.1. With the raw ODs
```{r}
plot_list <- list()
strains <- sort(unique(sample_ids$Sys.Name))
for (i in 1:length(strains)) {
  strain <- strains[i]
  strain_type <- sample_ids$Status[sample_ids$Sys.Name == strain][1]
  gene_name <- sample_ids$Gene.name[sample_ids$Sys.Name == strain][1]
  temp <- growth_long_full_raw %>%
    dplyr::filter(Strain == strain)
  if (strain_type == "Synthetase") {
    plot_list[[i]] <- ggplot(data = temp, aes(x = Time, y = OD_raw, group = as.factor(Temp_set), col = as.factor(Temp_set))) +
    geom_line(linewidth = 1) +
    theme_light() +
    xlab("Time (s)") +
    ylab("OD 600") +
    labs(title = glue("Growth curves for strain {strain} - {strain_type}"),
         subtitle = glue("{gene_name}"),
         col = "Temperature (ºC)") +
      coord_cartesian(ylim = c(0, 1.25))
  } else 
  if (strain_type == "Lethal") {
    lethal_temp <- sample_ids$Lethal_temp[sample_ids$Sys.Name == strain][1]
    plot_list[[i]] <- ggplot(data = temp, aes(x = Time, y = OD_raw, group = as.factor(Temp_set), col = as.factor(Temp_set))) +
    geom_line(linewidth = 1) +
    theme_light() +
    xlab("Time (s)") +
    ylab("OD 600") +
    labs(title = glue("Growth curves for strain {strain} - {strain_type}"),
         subtitle = glue("{lethal_temp}"),
         col = "Temperature (ºC)") +
      coord_cartesian(ylim = c(0, 1.25))
  }
  else
  {
    plot_list[[i]] <- ggplot(data = temp, aes(x = Time, y = OD_raw, group = as.factor(Temp_set), col = as.factor(Temp_set))) +
    geom_line(linewidth = 1) +
    theme_light() +
    xlab("Time (s)") +
    ylab("OD 600") +
    labs(title = glue("Growth curves for strain {strain} - {strain_type}"),
         col = "Temperature (ºC)") +
      coord_cartesian(ylim = c(0, 1.25))
  }
}

## List of ggplot objects has to be turned to list of grobs for grid.arrange to work
my_grobs <- lapply(plot_list, ggplotGrob)

## Print to PDF
output_file = paste(base_dir, "TSSC/Output/Plots/pilot/growth_curves/gcs_trna_synthetases_raw_", which_try, ".pdf", sep = "")
ggsave(
   filename = output_file, 
   plot = marrangeGrob(my_grobs, nrow=3, ncol=3), 
   width = 15, height = 9
)
```

## 3.2. With the ODs normalized to time = 0
```{r}
plot_list <- list()
temp <- sample_ids %>%
  dplyr::arrange(Status)
strains <- unique(temp$Sys.Name)
for (i in 1:length(strains)) {
  strain <- strains[i]
  strain_type <- sample_ids$Status[sample_ids$Sys.Name == strain][1]
  gene_name <- sample_ids$Gene.name[sample_ids$Sys.Name == strain][1]
  temp <- growth_long_full_norm %>%
    dplyr::filter(Strain == strain)
  if (strain_type == "Synthetase") {
    plot_list[[i]] <- ggplot(data = temp, aes(x = Time, y = OD_norm, group = as.factor(Temp_set), col = as.factor(Temp_set))) +
    geom_line(linewidth = 1) +
    theme_light() +
    xlab("Time (s)") +
    ylab("Ratio to OD at T0") +
    labs(title = glue("Growth curves for strain {strain} - {strain_type}"),
         subtitle = glue("{gene_name}"),
         col = "Temperature (ºC)") +
      coord_cartesian(ylim = c(0, 12.5))
  } else 
  if (strain_type == "Lethal") {
    lethal_temp <- sample_ids$Lethal_temp[sample_ids$Sys.Name == strain][1]
    plot_list[[i]] <- ggplot(data = temp, aes(x = Time, y = OD_norm, group = as.factor(Temp_set), col = as.factor(Temp_set))) +
    geom_line(linewidth = 1) +
    theme_light() +
    xlab("Time (s)") +
    ylab("OD 600") +
    labs(title = glue("Growth curves for strain {strain} - {strain_type}"),
         subtitle = glue("{lethal_temp}"),
         col = "Temperature (ºC)") +
      coord_cartesian(ylim = c(0, 12.5))
  }
  else
  {
    plot_list[[i]] <- ggplot(data = temp, aes(x = Time, y = OD_norm, group = as.factor(Temp_set), col = as.factor(Temp_set))) +
    geom_line(linewidth = 1) +
    theme_light() +
    xlab("Time (s)") +
    ylab("Ratio to OD at T0") +
    labs(title = glue("Growth curves for strain {strain} - {strain_type}"),
         col = "Temperature (ºC)") +
      coord_cartesian(ylim = c(0, 12.5))
  }
}

## List of ggplot objects has to be turned to list of grobs for grid.arrange to work
my_grobs <- lapply(plot_list, ggplotGrob)

## Print to PDF
output_file = paste(base_dir, "TSSC/Output/Plots/pilot/growth_curves/gcs_trna_synthetases_norm_to_t0_", which_try, ".pdf", sep = "")
ggsave(
   filename = output_file, 
   plot = marrangeGrob(my_grobs, nrow=3, ncol=3), 
   width = 15, height = 9
)
```




```{r}

```















